The Effect of Fictional Reappraisal on Subjective Ratings Toward Images

### 2025-11-13 paper/manuscript.qmd
minor

Recent advances in artificial intelligence (AI) have introduced new challenges for human cognition, particularly regarding the ability to distinguish between authentic and fabricated experiences [@miller2023ai]. 
The stakes of such perceptual uncertainty are considerable, with misinformation representing one of the most pressing societal consequences [@kreps2022all].
For instance, deepfakes (face-swapping technologies that enable the creation of realistic fake images and videos) have already been employed to fabricate convincing political speeches that appear genuine [@meskys2020regulating].
When deployed in electoral contexts, such content has the potential to distort public opinion and undermine democratic processes by rendering the truth unclear [@graber2021artificial]. 
Yet political disinformation is only one domain in which synthetic media exerts influence. 
The increased accessibility of generative technologies means that manipulated or entirely artificial content now permeates social media, entertainment, immersive environments, and interpersonal communication [@nightingale2022ai].  
In these settings, the boundaries between the “real” and the “artificial” are becoming progressively less discernible, raising fundamental questions about how individuals perceive and evaluate reality.

@@

A key challenge lies in the prevalence of ambiguous stimuli, that is images, texts, videos, or environments whose authenticity cannot be easily established. 
While artificial stimuli once carried perceptual markers that betrayed their inauthenticity, such as distortions in early computer-generated imagery [@corvi2023detection; @mcdonnell2010face], these limitations are rapidly disappearing. 
In the domain of face generation especially, current algorithms now produce synthetic images that are virtually indistinguishable from genuine photographs [@moshel2022you; @nightingale2022ai]. 
As these technologies approach perfect realism, the study of reality perception becomes not only a theoretical concern but also a practical imperative, with direct implications for information security, social trust, and the ethical deployment of AI.

<!-- Affective Responses and Reality Judgments
  - Link perception of realism with emotional engagement: arousal, valence, enticement.
  - Mention theories such as Affective Reality Theory (Makowski, 2023) if relevant, or more general emotion–cognition links.
  - Emotions may not just follow perception, but also shape how we judge what is real. -->

>>>>

<!-- Duffy et al., 2024:  when image sare controlled for types of artifacts participants show no ability in distinguishing real vs fake faces. Altough, they did rate faces with higher intensity smiles as more real. -->

Recent advances in artificial intelligence (AI) have introduced new challenges for human cognition, particularly regarding the ability to distinguish between authentic and fabricated experiences [@miller2023ai].

The consequences of such perceptual uncertainty are considerable, with misinformation representing one of the most pressing societal consequences [@kreps2022all].
For instance, deepfakes (face-swapping technologies that enable the creation of realistic fake images and videos) have already been employed to fabricate convincing political speeches that appear genuine [@meskys2020regulating].
When deployed in electoral contexts, such content has the potential to distort public opinion and undermine democratic processes by rendering the truth unclear [@graber2021artificial]. 
Yet political disinformation is only one domain in which synthetic media exerts influence. 
The increased accessibility of generative technologies means that manipulated or entirely artificial content now permeates social media, entertainment, immersive environments, and interpersonal communication [@nightingale2022ai].  
In these settings, the boundaries between the “real” and the “artificial” are becoming progressively less discernible, raising fundamental questions about how individuals perceive and evaluate reality.

@@

A key challenge lies in the prevalence of ambiguous stimuli, that is images, texts, videos, or environments whose authenticity cannot be easily established. 
While artificial stimuli once carried perceptual markers that made their inauthenticity easier to spot, such as distortions in early computer-generated imagery [@corvi2023detection; @mcdonnell2010face], these limitations are rapidly disappearing. 
In the domain of face generation especially, current algorithms now produce synthetic images that are virtually indistinguishable from genuine photographs [@moshel2022you; @nightingale2022ai]. 
As these technologies approach perfect realism, the study of reality perception becomes not only a theoretical concern but also a practical imperative, with direct implications for information security, social trust, and the ethical deployment of AI.

<!-- Affective Responses and Reality Judgments
  - Link perception of realism with emotional engagement: arousal, valence, enticement.
  - Mention theories such as Affective Reality Theory (Makowski, 2023) if relevant, or more general emotion–cognition links.
  - Emotions may not just follow perception, but also shape how we judge what is real -PP. -->
